---
title: "Lab 6 Comparing two means"
author: "Hamze Omar"
date: 23/04/2020
output: github_document
---

```{r}

# load packages -----------------------------------------------------------

library(tidyverse)

# read data ---------------------------------------------------------------

fish <- read_csv("chap12q19ElectricFish(1).csv")
```



```{r}
fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) %>% 
  print()

fish_long_data <- fish_long %>% 
  group_by(location) %>% 
  summarize(
    n = n(),
    mean = mean(species),
    sd = sd(species),
    sem = sd/sqrt(n),
    upper = mean + 1.96 * sem,
    lower = mean - 1.96 * sem
  ) %>% 
  print()
```


Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

## Question A

> What is the mean different in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

## ANSWER

I will write my calculation of the mean differences below.

# the mean difference of upstream and downstream

downstream mean = 16.41667, upstream mean = 14.58333, difference = 1.83334

# 95% confidence interval

UCI = 8.253697
LCI = -4.587031


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
t.test(formula = species ~ location, data = fish_long)
```

## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

# ANSWER

The P-value indicates that it's greater than the alpha level of 0.05, in that case we failed to reject the null hypothesis. P-value is 0.5596.....

```{r fish_long, message=FALSE, warning=FALSE, paged.print=FALSE}
t.test(formula = species ~ location, data = fish_long)

fish_long %>% 
    ggplot(aes(x = location, y = species)) +
    geom_jitter(aes(color = location), 
                shape = 16, size = 3, 
                alpha = 0.3, width = 0.4) +
    geom_errorbar(aes(y = mean, ymax = upper, ymin = lower), 
                  data = fish_long_data, 
                  width = .1, size = .8) +
    geom_point(aes(y = mean), 
               data = fish_long_data, 
               size = 3) +
    scale_color_manual(values = c("dark blue","red")) +
    theme_minimal() +
    guides(color = "none")
```

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

# ANSWER

The distribution is not  normally distributed and the means of both variables show some differences. From the previous calculation, it tells us that the true mean difference isn't equal to zero.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
fish_long %>% 
    ggplot(aes(x = species)) +
    geom_histogram(
      aes(fill = location), 
      bins = 20, 
      alpha = 0.5, 
      position = "identity"
    ) +
    scale_fill_manual(values = c("darkorange","cyan4")) +
    theme_minimal()
```



```{r}
crab_fans_data <- read_csv("chap15q27FiddlerCrabFans.csv") %>%
    rename(type = crabType, temperature = bodyTemperature)
  crab_fans_data
```

## ANOVA

Fiddler crabs are so called because males have a greatly enlarged "major" claw, which is used to attract females and to defend a burrow.

Darnell and Munguia (2011) recently suggested that this appendage might also act as a heat sink, keeping males cooler while out of the burrow on hot days.

To test this, they placed four groups of crabs into separate plastic cups and supplied a source of radiant heat (60-watt light bulb) from above. The four groups were intact male crabs, male crabs with the major claw removed; male crabs with the other (minor) claw removed (control); and intact female fiddler crabs.

They measured the body temperature of crabs every 10 minutes for 1.5 hours. These measurements were used to calculate a rate of heat gain for every individual crab in degrees C/log minute. Rates of heat gain for all crabs are provided in the accompanying data file.

### Question D

Graph the distribution of body temperatures for each crab type:

```{r}

temperature_means <-
    crab_fans_data %>%
    filter(!is.na(temperature)) %>%      # remove missing values
    group_by(type) %>%
    summarize(
      mean = mean(temperature),
      sd = sd(temperature),
      n = n(),
      sem = sd / sqrt(n),
      upper = mean + 1.96 * sem,
      lower = mean - 1.96 * sem)


 ggplot(data = crab_fans_data, aes(x = type, y = temperature)) +
    geom_jitter(aes(color = type),
                width = 0.1,
                alpha = 0.7,
                show.legend = FALSE,
                na.rm = TRUE) +
    geom_errorbar(aes(y = mean, ymin = lower, ymax = upper), 
                  data = temperature_means,
                  width = .1, position = position_nudge(.3)) +
    geom_point(aes(y = mean), data = temperature_means,
               position = position_nudge(.3)) +
    scale_color_manual(values = c("darkorange","darkorchid","cyan4", "red"))
```

### Question E

Does body temperature varies among crab types? State the null and alternative hypothesis, conduct and ANOVA, and interpret the results.

Yes, body temp varies in different scenerios, for example if we look at the female crab, they have the highest body temp. In addition if we compare intact male, male major, and minor removed.
Intact male shows higher temp of heat, but if we subset from male major, the temp decreases. On the minor male removed the heat temp still as high but skuad to the left of the graph

```{r}
 crab_fans_data %>% filter(!is.na(type)) %>% 
    ggplot(aes(x = temperature)) +
    geom_histogram(
      aes(fill = type), 
      bins = 10, 
      alpha = 0.5, 
      position = "identity", na.rm = TRUE
    ) + scale_fill_manual(values = c("darkorange","cyan4", "darkorchid", "red")) +
    theme_minimal() +
    facet_wrap(~type, ncol=1)

aov_bill_depth_species <-
  aov(temperature ~ type, data = crab_fans_data)
aov_bill_depth_species

summary(aov_bill_depth_species)
```