---
title: "Lab 6 Comparing two means"
author: "Hamze Omar"
date: 23/04/2020
output: github_document
---

```{r}

# load packages -----------------------------------------------------------

library(tidyverse)

# read data ---------------------------------------------------------------

fish <- read_csv("chap12q19ElectricFish(1).csv")
```

```{r}
fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) %>% 
  print()

fish_long_data <- fish_long %>% 
  group_by(location) %>% 
  summarize(
    n = n(),
    mean = mean(species),
    sd = sd(species),
    sem = sd/sqrt(n),
    upper = mean + 1.96 * sem,
    lower = mean - 1.96 * sem
  ) %>% 
  print()
```


Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

## Question A

> What is the mean different in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

## ANSWER

I will write my calculation of the mean differences below.

# the mean difference of upstream and downstream

downstream mean = 16.41667, upstream mean = 14.58333, difference = 1.83334

# 95% confidence interval

UCI = 8.253697
LCI = -4.587031


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
t.test(formula = species ~ location, data = fish_long)
```

## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

# ANSWER

The P-value indicates that it's greater than the alpha level of 0.05, in that case we failed to reject the null hypothesis. P-value is 0.5596, this means there's 55%.

```{r fish_long, message=FALSE, warning=FALSE, paged.print=FALSE}
t.test(formula = species ~ location, data = fish_long)
```

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

# ANSWER

The distribution is normally distributed and the means of both variables show some differences, the true mean difference isn't equal to zero.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
fish_long %>% 
    ggplot(aes(x = location, y = species)) +
    geom_jitter(aes(color = location), 
                shape = 16, size = 3, 
                alpha = 0.3, width = 0.4) +
    geom_errorbar(aes(y = mean, ymax = upper, ymin = lower), 
                  data = fish_long_data, 
                  width = .1, size = .8) +
    geom_point(aes(y = mean), 
               data = fish_long_data, 
               size = 3) +
    scale_color_manual(values = c("dark blue","red")) +
    theme_minimal() +
    guides(color = "none")
```
